Secret-Key Based Public-Key Encryption

### 2017-05-26 latex/epq/paper.tex
added docstrings

\begin{abstract}
  We define an algorithm for performing a key exchange between two parties utilizing lattice based ideas. The algorithm makes use of a small number of bignum multiplications and additions and is conceptually simple.\\ 
\end{abstract}

@@

The introduction of Quantum Computing and post-quantum cryptography has lead to the development of lattice based cryptographic schemes that are believed to be secure even when faced with quantum parallelism. Simply put, RSA and DH will need replacement at some point, and lattices seem like a promising alternative.

Lattices are also very attractive because they offer homomorphic properties, which is another of the most important topics in cryptography at the moment. This property is also extremely important because it allows for the construction of simple public key encryption schemes and key exchange shemes from secret key ciphers. We present such a scheme in this paper, as well as a homomorphic cipher suitable for instantiating the scheme.

>>>>

\begin{abstract}
  We define algorithms for performing public key encryption and key exchange based on secret key ciphers that offer a homomorphic property. The nature of the design(s) allows for instantiation with any secure homomorphic cipher. The public keys in the presented schemes are amenable to being randomized, which offers the prospect of encrypted/anonymous public keys.
\end{abstract}

@@

Two of the most important topics in cryptography right now are: replacing RSA and DH in practice, and exploring the potential of homomorphic encryption. In this work we examine the application of homomorphic encryption towards the creation of simple public key encryption and key exchange algorithms.

Some of the algorithms are not necessarily novel, but are more explicit and general descriptions of previous public key encryption algorithms; For example, the original Merkle-Hellman Knapsack Cryptosystem is effectively an example of a secret key homorphic cipher being used to create a public key, while inverting the public key encryption function is equivalent to performing a subset sum problem. We generalize the method and encodings for these techniques in this work.

We will examine these systems from several perspectives: As the subset sum problem, and as a lattice problem.

### 2017-05-26 latex/secretkeybasedpublickeyencryption/paper.tex
added docstrings

\author{Anonymous\inst{1}}
\institute{City, State \email{address@provider.com}}
\title[\texttt Design date]{\texttt Design date}

@@

% use optional argument because the \LaTeX command breaks the PDF keywords
\keywords[Key Exchange]{Key Exchange}\todo{Keywords?}

\begin{abstract}
  We define an algorithm for performing a key exchange between two parties utilizing lattice based ideas. The algorithm makes use of a small number of bignum multiplications and additions and is conceptually simple.\\ 
\end{abstract}

@@

\section{Introduction}
 The ability to exchange cryptographic keys over insecure channels is of vital importance to the function of the internet; Without algorithms for arranging secure communication online, much of every day life in the modern world would cease to be possible.\\

The introduction of Quantum Computing and post-quantum cryptography has lead to the development of lattice based cryptographic schemes that are believed to be secure even when faced with quantum parallelism. Simply put, RSA and DH will need replacement at some point, and lattices seem like a promising alternative.

Lattices are also very attractive because they offer homomorphic properties, which is another of the most important topics in cryptography at the moment. This property is also extremely important because it allows for the construction of simple public key encryption schemes and key exchange shemes from secret key ciphers. We present such a scheme in this paper, as well as a homomorphic cipher suitable for instantiating the scheme.

\section{Definitions}
We use $+$ to denote addition, $ab$ to denote multiplication, and $\mod p$ to denote the modulo operation.

\section{Algorithm}
First, we shall define the key exchange algorithm; Then, we shall define a secret key cipher suitable for instantiating the key exchange algorithm.

\subsection{Key Exchange Algorithm}
Let $pb_1$ and $pb_2$ be two encryptions of $0$ under the homomorphic secret key cipher. These two values constitute a public key. To share a secret with the owner of the public key, compute:
\begin{align}
    e \gets secret \\
    c \gets (pb_1  r_1) + (pb_2 r_2) + e
\end{align}

Then the value $c$ is shared with the owner of the public key. The owner of the public key can recover $e$ by applying the homomorphic decryption operation of the secret key cipher to c; This is because $pb_1$ is an encryption of $0$, $pb_1 r_1$ is equivalent to $0 r_1 \equiv 0$, similar for $pb_2$. So when the decryption operation is applied, the result is $0 + 0 + e \equiv e$.

\subsection{Secret Key Cipher}
Creation of the $pb_1$ and $pb_2$ values requires a secure, additively homomorphic cipher. We define a simple lattice based cipher which utilizes effectively the same operations as the key exchange process. Let $p_1$ and $p_2$ be two large numbers, where $log_2(p_1) >= log_2(p_2) * 3$; These two numbers constitute the secret key for the cipher, which in turn acts as the private key for the key exchange presented previously.\\

To encrypt a ciphertext, compute:

\begin{align}
   c \gets (p_1 r_1) + (p_2 r_2) + m
\end{align}

Where the $r$ values are of appropriate size relative to $p_1$ and $p_2$ \todo{what sizes are appropriate}
While decryption is computed as:

\begin{align}
    m \gets (c \mod p_1) \mod p_2
\end{align}

\subsection{Note}
While the decryption circuit works via two modulo operations of the base points of the lattice, this operation does not work for the public key/key exchange portion of the algorithm - The reason is because in the secret key cipher, the two different lattice points are of largely different size; In the key exchange algorithm, both points are roughly the same size, and so the double modulo operation does not function identically.

\section{Conclusion}
 We define a simple lattice based key exchange algorithm. The algorithm utilizes 2 bignum multiplications and 2 bignum additions to generate a keypair, 2 bignum multiplications and 2 bignum additions per key sent, and 2 modulo operations on bignums to receive a key. Public keys typically occupy < 1200 bits using current parameter cohices, while private keys require about 600 bits.
\end{document}

>>>>

\author{Ella Rose\inst{1}}
\institute{Paso Robles, CA\email{python_pride@protonmail.com}}
\title[Secret-Key Based Public-Key Encryption]{Secret-Key Based Public-Key Encryption}

@@

% use optional argument because the \LaTeX command breaks the PDF keywords
\keywords[Homomorphic Encryption, Symmetric Encryption, Asymmetric Encryption]{Homomorphic Encryption, Symmetric Encryption, Asymmetric Encryption}

\begin{abstract}
 We present a perspective which demonstrates not only that secret-key based public key encryption is possible, but that well known primitives already fit into this framework. We describe a variety of such constructions, which are of varying complexity and security.
\end{abstract}

@@

\section{Introduction}
 Ciphers that support homomorphic operations on ciphertexts are very attractive because they tend to make the creation of public key encryption schemes relatively straightforward and simple. This paper attempts to illustrate and bring to light this fact. We will begin by covering "classic" cryptosystems such as Merkle's original knapsack cryptosystem, as well as modular exponentiation based cryptosystems such as DH and RSA.

\section{Definitions}
We use $+$ to denote addition, $ab$ to denote multiplication, $a / b$ to denote division, and $\mod p$ to denote the modulo operation. Exponentiation is denoted as $a \pow b$.

\section{Merkle's Knapsack Cryptosystem}
\subsection{Basic Concept}
This original cryptosystem is actually slightly more complex then some of the others that will be described. This is largely due to the fact that decryption involves actually solving a subset sum problem.

The secret key cipher in this cryptosystem is modular multiplication. The plaintext numbers, as well as the modulus, are kept secret, in addition to the key.

To create a public key, the secret key cipher is used to encrypt a set of numbers. The numbers are chosen such that utilizing the set to perform a subset sum problem will result in an easy problem to solve, allowing recovery of the summed elements. After the numbers have been encrypted, the set of numbers will be randomized and using them in a subset sum will result in greater difficulty inverting the sum.

The public key encryption process involves summing elements of the public key according to the bits of the message to be sent. Decrypting a ciphertext produced this way is equivalent to solving a subset sum problem.

The secret key is used to turn the summed elements back into the "easy version" of the numbers, so that the subset sum problem can then be solved.

\subsection{Key Generation}
\subsubsection{Encryption}
\subsubsection{Decryption}

\subsection{Randomization}
Any homomorphic cipher requires randomization in order to be secure. The presented design is very simple to randomize; Continuing with the previous example, we would simply generate an additional $k_3$ and a random $r_0$ and include that in the sum as well:

\section{Conclusion}

\end{document}
